Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
pf <- read.csv('pseudo_facebook.tsv', sep = "\t")
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)

pfplot <-qplot(data = pf, x=dob_day) +
  scale_x_continuous(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response: too many bdays on 1st

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
qplot(data = pf, x=dob_day) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month, ncol = 3)


pfplot +
  facet_grid(. ~ dob_month)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
names(pf)
pffriend <- qplot(data = pf, x=friend_count)
pffriend
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes: 2 ways of doing the limits

```{r Limiting the Axes}
qplot(data = pf, x=friend_count, xlim = c(0,1000))

qplot(data = pf, x=friend_count) +
  scale_x_continuous(limits = c(0,1000))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender,ncol = 2)
```

***

### Omitting NA Values
Notes: 2 ways

```{r Omitting NA Values}
# na.omit() removes nas from all variables
qplot(x = friend_count, data = na.omit(pf), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender,ncol = 2)

#subset() removes nas from specified variable
qplot(x = friend_count, data = subset (pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender,ncol = 2)


```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)

by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response: women

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response: median more robust, it's 50%, mean is pulled by extreme values

***

### Tenure
Notes:

```{r Tenure}
qplot(data = pf, x = tenure, binwidth = 30,
      color = I('black'), fill = I('#099DD9'))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(data = pf, x = tenure/365, binwidth = .25,
      color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(breaks = (1:7), limits = c(0,7))

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(data = pf, x = tenure/365, binwidth = .25,
      xlab = 'label for x axis',
      ylab = 'another label',
      color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(breaks = (1:7), limits = c(0,7))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(data = pf, x = age,  binwidth=1,
      xlab = 'label for x axis',
      ylab = 'another label',
      color = I('black'), fill = I('#099DD9'))
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
#adding the log scale
library(gridExtra)
p1 <- qplot(data = pf, x=friend_count,color = I('black'), fill = I('#099DD9')) 
p2 <- qplot(data = pf, x=log10(friend_count),color = I('black'), fill = I('#099DD9')) 
p3 <- qplot(data = pf, x=sqrt(friend_count),color = I('black'), fill = I('#099DD9')) 
#grid.arrange comes from gridExtra
grid.arrange(p1,p2,p3,ncol=1)



p1 <- ggplot(aes(x=friend_count), data = pf) + 
  geom_histogram()
P2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()
# also works with qplot and scale_x_command

p4 <- qplot(data = pf, x=log10(friend_count))
p5 <- qplot(data = pf, x=friend_count) + scale_x_log10()
grid.arrange(p4,p5, ncol = 2)
#note different scales!!!

rm(p1,p2,p3,p4,p5)

```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(data = subset(pf,!is.na(gender) ), x=www_likes, geom = 'freqpoly', color=gender) + scale_x_continuous() +
  scale_x_log10()
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
summary(pf$www_likes)
by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(x = gender, y = friend_count, data = subset (pf, !is.na(gender)), geom ='boxplot' ) + scale_y_log10()
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x = gender, y = friend_count, data = subset (pf, !is.na(gender)), geom ='boxplot' ) +
  coord_cartesian(ylim = c(0,1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(x = gender, y = friend_count, data = subset (pf, !is.na(gender)), geom ='boxplot' ) +
  coord_cartesian(ylim = c(0,250))

by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
qplot(x=gender, y=friendships_initiated, data=subset(pf, !is.na(gender)), geom ='boxplot' ) +
  coord_cartesian(ylim = c(0,150))

by(pf$friendships_initiated, pf$gender, summary)
```

Response:

***

### Getting Logical
Notes: if there are many 0s in data and we just want to check presence/absence (value>0)

```{r Getting Logical}
summary(pf$mobile_likes)
summary(pf$mobile_likes>0)

pf$mobile_checkin <- ifelse(pf$mobile_likes>0, 1,0)
#convert variable to factor
pf$mobile_checkin <- factor(pf$mobile_checkin)
summary(pf$mobile_checkin)

#calculate percentage of users that check in with mobile
#sum of all values with 1 divided by count (length) of all records
sum(pf$mobile_checkin==1)/length(pf$mobile_checkin)

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!