Lesson 5
========================================================
```
{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable} 
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
library(ggplot2)
library(dplyr)
library(alr3)

ggplot(data=subset(pf, !is.na(gender)),aes(gender,age))  +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom='point',shape=4)

ggplot(data=subset(pf, !is.na(gender)),aes(age,friend_count))+
  geom_line(aes(color=gender), stat = 'summary', fun.y=median)


pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>% # one grouping will be removed by <summarise>, need to remove second
  arrange(age)

head(pf.fc_by_age_gender)
```
```{r}
ggplot(data=subset(pf, !is.na(gender)),aes(age,friend_count))+
  geom_line(aes(color=gender), stat = 'summary', fun.y=median)
```
```{r}
library(dplyr)
#chaining functions together
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>% # one grouping will be removed by <summarise>, need to remove second
  arrange(age)

head(pf.fc_by_age_gender)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(data=pf.fc_by_age_gender, aes(age,median_friend_count))+
  geom_line(aes(color=gender))
```

***

### Thinking in Ratios
Notes: 

***

### Wide and Long Format
Notes: take repeating values from one column and use them as separate columns (e.g. gender - male/female)
can be done with reshape2 or with tidyr

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender_wide <-dcast(pf.fc_by_age_gender,
                                 age ~ gender,
                                 value.var = 'median_friend_count')

head(pf.fc_by_age_gender_wide)
melt(pf.fc_by_age_gender_wide)

```

```{r reshape with tidyr}
library(tidyr)

spread(subset(pf.fc_by_age_gender,
              select = c('gender','age','median_friend_count')),
       gender,median_friend_count)
```

***

### Ratio Plot
Notes:plot ratio of female to male meddian friend counts with new dataframe,
include horisontal line at intercept = 1, make it dashed

```{r Ratio Plot}
ggplot(data=pf.fc_by_age_gender_wide,aes(age, female/male))+
  geom_line()+
  geom_hline(yintercept = 1, linetype=2, alpha=.7) #horizontal line
```

***

### Third Quantitative Variable
Notes: create var year joined, using tenure (days), and 2014 as reference year

```{r Third Quantitative Variable}

pf$year_joined <- floor(2014 - pf$tenure/365)

```

***

### Cut a Variable
Notes: cut the variable year joined into buckets by age (group couple years together)

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)
```
```{r}
?cut
pf$year_joined_bucket <-cut(pf$year_joined,
                            breaks = c(2004, 2009, 2011, 2012,2014)
                            )

```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(data=subset(pf, !is.na(pf$year_joined_bucket)),
       aes(age,friend_count))+
  geom_line(aes(color=year_joined_bucket),
            stat = 'summary',
            fun.y=median)
```
Users with longer tenure have more friends, with exception of old ones
***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(data=subset(pf, !is.na(pf$year_joined_bucket)),
       aes(age,friend_count))+
  geom_line(aes(color=year_joined_bucket), 
            stat = 'summary',
            fun.y=mean)+ #means for separate buckets
  geom_line(stat = 'summary', fun.y = 'mean', linetype=2) #grand mean (all data)
```

***

### Friending Rate
Notes: find friend_count / tenure ratio, median, mean, for tenure >0

```{r Friending Rate}
with(subset(pf, tenure>0), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes: plot friendships initiated/day vs. tenure
use: age, tenure, friendships_initiated and year_joined, bucket)

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
ggplot(data=subset(pf, tenure>0),
       aes(tenure,friendships_initiated/tenure))+
  geom_line(aes(color=year_joined_bucket), 
            stat = 'summary',
            fun.y=mean)


#make it smooth:
ggplot(data=subset(pf, tenure>0),
       aes(tenure,friendships_initiated/tenure))+
  geom_smooth(aes(color=year_joined_bucket))

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined_bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)
yo$id <- factor(yo$id)
str(yo)

install.packages("tidyr")
library(ggplot2)

ggplot(data=yo, aes(x=price))+
  geom_histogram()

qplot(data = yo, x=price, color = I('black'), fill = I('#099DD9'))
```
```{r}
head(yo)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo)
#number of unique prices
length(unique(yo$price))
#let's list the unique prices
table(yo$price)

#we know separate purchse/flavor, per household, at given date, create variable that combines values from several columns (flavors of yogurth)
#use transform
?transform
yo <- transform(yo, all.purchases = strawberry+blueberry+pina.colada+plain+mixed.berry)
 str(yo)
#alternate verbose version
 yo$all.purchases <- yo$strawberry + yo$blueberry + yo$pina.colada + yo$plain + yo$mixed.berry


```

***

### Prices over Time
Notes:

```{r Prices over Time}
qplot(data = yo, x=all.purchases, color = I('black'), fill = I('#099DD9'), binwidth=1)
```
```{r}
#price over time
ggplot(yo,aes(time,price))+
  geom_point(alpha=.3, shape=21)
```

***

### Sampling Observations
Notes: if working with multiple observations of same units work with some samples of those units, to see e.g. between household variation

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
?set.seed
set.seed(4230)

sample.ids <- sample(levels(yo$id), 16)
sample.ids
```
```{r}
ggplot(subset(yo, id %in% sample.ids), #the way to subset the data using another set, like join in sql
       aes(time,price))+
  facet_wrap( ~ id)+
  geom_line()+
  geom_point(aes(size=all.purchases), pch=1)
```

***

### The Limits of Cross Sectional Data
Notes: we were faceting over households with yo data


***

### Many Variables
Notes: use to discover relationships between variables we were not interested in

***

### Scatterplot Matrix
Notes: 
```{r}
install.packages("GGally")
library(GGally)
theme_set(theme_minimal(20))

set.seed(1836)
pf_subset <-pf[,c(2:15)]
names(pf_subset)
```
```{r}
#this will take plenty time to run
ggpairs(pf_subset[sample.int(nrow(pf_subset),1000), ])
#explore the ggpairs syntax

#if too much, one can subset:
pf_subset = pf[, c('age', 'dob_year', 'dob_month', 'gender', 'tenure')]
#OR
pf_subset <- subset(pf, select = -c(userid, year_joined, year_joined_bucket))
#The - sign in the "select" value indicates all but the listed columns.

```


***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)
```
```{r heatmap}
ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

