Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(data = pf, x = age, y = friend_count)

#don't need to specify axes: (first x, second y)
qplot(data = pf, age, friend_count)


```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(age,friend_count), data = pf) + geom_point()

summary(pf$age)

ggplot(aes(age,friend_count), data = pf) + geom_point(alpha=.1) +
  xlim(13,90)


```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(age,friend_count), data = pf) + geom_jitter(alpha = .1, size = .2) +
  xlim(13,90) 
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(age,friend_count), data = pf) + geom_point(alpha=.1, size=.1, position = position_jitter(h=0)) +
  xlim(13,90) +
  coord_trans(y='sqrt') 
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(age,friend_count), data = pf) + geom_point(alpha=.1, size=.1, position = position_jitter(h=0)) +
  xlim(13,90) +
  coord_trans(y='sqrt') 
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(age,friendships_initiated), data = pf) + geom_point(alpha=.1, size=.1, position = position_jitter(h=0)) +
  xlim(13,90) +
  coord_trans(y='sqrt') 
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

age_groups <-group_by(pf,age)

head(age_groups)
pf.fc_by_age <-summarise(age_groups,
                         friendCountMean = mean(friend_count),
                         friendCountMedian = median(friend_count),
                         n=n())
pf.fc_by_age <-arrange(pf.fc_by_age, age)
head(pf.fc_by_age)


#chaining functions - use %>%
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friendCountMean = mean(friend_count),
            friendCountMedian = median(friend_count),
            n=n()) %>%
  arrange(age)
  head(pf.fc_by_age)
# this code creates variable pf.fc_by_age, from pf, which is grouped by age, then adds summaries: mean, median and n(count), arranges by age and prints out top 6 rows

```

Create your plot!

```{r Conditional Means Plot}
qplot(data=pf.fc_by_age, age,friendCountMean)

ggplot(data=pf.fc_by_age,aes(age,friendCountMean))+
  geom_point(alpha=.5) +
  geom_line(color='grey')
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(age,friend_count), data = pf)+
  geom_point(alpha=.1, size=.1, position = position_jitter(h=0), color = 'orange')+
  geom_line(stat = 'summary', fun.y = mean)+
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .1), linetype=2, color='blue')+
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .9), linetype=2, color='blue')+
  coord_cartesian(xlim = c(13,90))+
  scale_y_sqrt()
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count)

with(pf, cor.test(age,friend_count))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age<=70) , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes: Inferential/desctiprive statistics, read on it, read on correlation methods

***

## Create Scatterplots
Notes:

```{r}
ggplot(pf,aes(likes_received,www_likes_received))+
  geom_point(alpha=.2) +
  with(pf, xlim(0,quantile(likes_received,.95))) +
  with(pf, ylim(0,quantile(www_likes_received,.95))) +
  geom_smooth(method = 'lm', color = 'red', size=.5)
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
with(pf, cor.test(likes_received, www_likes_received))
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}
data("Mitchell")
?Mitchell
head(Mitchell)

str(Mitchell)

ggplot(Mitchell,aes(Month,Temp))+
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell, cor.test(Month, Temp))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(Mitchell,aes(Month,Temp))+
  geom_point()+
  scale_x_continuous(breaks = seq(0,203,12))+
  geom_line(color='grey')
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(data=pf.fc_by_age,aes(age,friendCountMean))+
  geom_point(alpha=.5) +
  geom_line(color='grey')
```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- with(pf, age + (1 - dob_month / 12))

```

Programming Assignment
```{r Programming Assignment}
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friendCountMean = mean(friend_count),
            friendCountMedian = median(friend_count),
            n=n()) %>%
  arrange(age_with_months)
  head(pf.fc_by_age)


```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
  ggplot(data=subset(pf.fc_by_age_months, age_with_months<71),aes(age_with_months,friendCountMean))+
  geom_line(color='grey')
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
  ggplot(data=subset(pf.fc_by_age_months, age_with_months<71),aes(age_with_months,friendCountMean))+
  geom_line(color='grey')+
  geom_smooth()
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

